Different rules for binocular combination of luminance flicker in cortical and subcortical pathways

### 2023-01-07 manuscript.tex
Added paragraph to discussion on blobs

\title{Different rules for binocular combination of luminance flicker in cortical and subcortical pathways}
\author{Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade \& Daniel H. Baker}
\date{2022-12-21}

@@

How does the brain combine information across the eyes? In primary visual cortex, spatial patterns are subject to `ocularity invariance', where monocular and binocular responses are rendered equal by mutual suppression between the eyes. Here we find a very different pattern of activity in response to luminance flicker, both in the cortex and also in the subcortical pathway that determines pupil diameter, by simultaneously collecting steady-state EEG and pupillometry data. We find strong binocular facilitation, and negligible interocular suppression from cortical sources, whereas pupil diameters showed weaker facilitation and stronger suppression. We confirm that this is not a consequence of the temporal parameters chosen, as pronounced binocular facilitation was also observed at faster flicker rates. Near-linear binocular combination was also found for the same stimuli using a matching task. A hierarchical Bayesian implementation of a standard binocular combination model confirms that interocular suppression was substantially weaker for our EEG and matching data, compared with the pupillometry results. These findings illustrate that the brain can repurpose a generic algorithm for different functions by adjusting parameters such as the weight of suppression between pathways.

@@

The brain must combine information across multiple sensory inputs to derive a coherent percept of the external world. This involves a process of signal combination both within (Baker and Wade, 2017) and between (Ernst and Banks, 2002) the senses. Binocular vision is a useful test case for signal combination, as the inputs to the two eyes overlap substantially (in species with forward-facing eyes), and the neural locus is well-established (Hubel and Wiesel, 1962). Yet much of our knowledge about binocular combination focuses on the `canonical' visual pathway, in which signals pass from the eyes to primary visual cortex (V1), via the lateral geniculate nucleus (LGN) (Purves et al., 2008). Signals are also combined across the eyes in a network of subcortical nuclei that govern pupil diameter in response to incoming light levels (McDougal and Gamlin, 2008). Our primary purpose here is to investigate whether signal combination involves the same or different operations across these two anatomically distinct pathways.

For pattern vision, binocular presentation confers greater sensitivity to low contrast targets than monocular presentation. This is known as binocular summation, with summation ratios (the relative improvement under binocular presentation) at detection threshold lying between \(\sqrt{2}\) and 2 (Baker et al., 2018; Campbell and Green, 1965). But this advantage is lost at high stimulus intensities, where both psychophysical performance (contrast discrimination thresholds) (Legge, 1984; Meese et al., 2006) and neural activity (Baker and Wade, 2017; Moradi and Heeger, 2009) are approximately equal for monocular and binocular presentation. Contemporary models of binocular vision (Ding and Sperling, 2006; Meese et al., 2006) advocate a process of interocular suppression that, at high contrasts, normalizes the two eyes inputs and negates the binocular advantage. This is consistent with our everyday experience of `ocularity invariance' (Baker et al., 2007): perceived contrast does not change when one eye is opened and closed.

The pupillary light reflex is an automatic constriction of the iris sphincter muscles in response to increases in light levels, which causes the pupil to shrink (McDougal and Gamlin, 2008). There is a clear binocular component to this reflex, as stimulation of one eye still causes constriction of the other eye's pupil (termed the consensual response; Wyatt and Musselman (1981)). Importantly, the neuroanatomical pathway involved completely bypasses the canonical cortical pathway (retina to V1), instead involving a network of subcortical nuclei, including the Pretectal Olivary nucleus, Superior Cervical ganglion, and Edinger-Westphal nucleus (Angée et al., 2021; Mathôt, 2018; McDougal and Gamlin, 2008; Wang and Munoz, 2015). To account for the consensual response, these brain regions must necessarily combine information from the left and right eyes (Doesschate and Alpern, 1967), yet the precise algorithm by which this is achieved is currently unclear.

To address this shortcoming, we designed an experiment that allowed us to simultaneously record electrophysiological and pupillometric responses to monocular and binocular stimuli. This novel paradigm allows us to probe both cortical (using EEG) and subcortical (using a binocular eyetracker) pathways simultaneously in response to flickering light, and make quantitative comparisons between them. Using periodic flicker entrains both neural (Norcia et al., 2015) and pupil (Spitschan et al., 2014) responses at the flicker frequency, enabling precise estimation of response amplitudes in the Fourier domain. We follow up our main experiment with additional exploration of the effect of stimulus frequency, and a psychophysical matching experiment measuring perceived flicker intensity (i.e.~temporal contrast). The results are interpreted using a hierarchical Bayesian computational model of binocular vision, and reveal that subcortical pathways involve stronger interocular suppression than the canonical cortical pathway.

@@

The stimuli were two discs of flickering light with a diameter of 3.74 degrees, presented on a black background. The same stimuli were used for all three experiments. Four dark red lines were added around both discs to help with their fusion into one binocular disc (see insert in Figure \ref{fig:pupildata}b for an example of the fused stimulus). The discs were viewed through a four-mirror stereoscope, which used front silvered mirrors to avoid internal reflections, and meant that participants saw a single fused disc. The use of a stereoscope allowed us to modulate the stimuli in three different ocular configurations: monocular, binocular, and dichoptic. Note that during monocular presentation of flicker, the unstimulated eye still saw the static (non-flickering) disc of mean luminance.

@@

The experiment was conducted in a windowless room, in which the only light source was the monitor. The participants sat at 99 cm from the monitor and the total optical viewing distance (through the stereoscope) was 107 cm. The experiment was carried out in one session lasting 45 minutes in total, divided in three blocks of 15 minutes each. In each block, there were 60 trials in total lasting 15 seconds each (12s of stimulus presentation, with an interstimulus interval of 3s). The participants were given no specific task instructions, but were asked to look at the fixation cross in the middle of the disc while trying to minimise their blinking during the presentation period.

We included six distinct ocular conditions, each at five temporal target contrast levels (combined factorially) relative to the mean luminance: 6, 12, 24, 48 and 96\%. Contrast is defined as temporal Michelson contrast; the difference between maximum and minimum luminances, scaled by the mean and expressed as a percentage. In the first three conditions, the discs flickered at 2 Hz, in either a monocular, binocular, or dichoptic arrangement. In the dichoptic condition the non-target eye saw a fixed contrast of 48\%. In the remaining three conditions (termed the cross-frequency conditions) one eye's disc flickered at 1.6Hz, and the other eye's disc flickered at 2Hz. We also tested monocular (one eye sees 1.6Hz flicker, the other sees mean luminance), binocular (one eye sees each frequency at the target contrast) and dichoptic (target stimulus flickering at 2Hz, mask contrast of 48\% at 1.6Hz in the other eye) arrangements. The rationale for flickering both eyes at 2Hz is that we can then measure summation behaviour between the eyes in the pupil and EEG response at 2Hz. The rationale for flickering the eyes at different frequencies is that this permits measurement of suppression between the eyes (i.e.~the reduction in the 2Hz response when a 1.6Hz mask component is added to the other eye). We counterbalanced presentation of the target stimulus across the left and right eyes.

@@

EEG data were converted from the ANT-EEProbe format to a compressed csv text file using a custom Matlab script and components of the EEGlab toolbox (Delorme and Makeig, 2004). The data for each participant were then loaded into R for analysis, where a ten-second waveform for each trial at each electrode was extracted (omitting the first two seconds). The fast Fourier transform was calculated for each waveform, and the spectrum stored in a matrix. All repetitions of each condition were then averaged for each electrode. They were then averaged across four occipital electrodes (\emph{POz}, \emph{Oz}, \emph{O1}, \emph{O2}), to obtain individual results. Finally, these were averaged across participants to obtain the group results. All averaging retained the phase information (i.e.~coherent averaging), and at each stage we excluded data points with a Mahalanobis distance exceeding \(D\) = 3 from the complex-valued mean (see Baker, 2021). For statistical comparisons of complex-valued data, we use the \(ANOVA^2_{circ}\) statistic described by Baker (2021). This is a multivariate extension of ANOVA that assumes equal variance of the real and imaginary Fourier components, or equivalently, an extension of the \(T^2_{circ}\) statistic of Victor and Mast (1991) that can compare more than two conditions.

A similar analysis pipeline was adopted for the pupillometry data. The data were converted from mp4 videos to a csv text file using the Pupil Player software (Kassner et al., 2014), which estimated pupil diameter for each eye on each frame using a 3D model of the eyeball. The individual data were then loaded into R for analysis, where again a ten-second waveform for each trial in each eye was extracted (excluding the first two seconds after stimulus onset). We interpolated across any dropped or missing frames to ensure regular and continuous sampling over time. The fast Fourier transform was calculated for each waveform, and all repetitions of each condition were pooled across eye and then averaged. Finally, data were averaged across all participants to obtain the group results. Again, we used coherent averaging, and excluded outlying data points in the same way as for the EEG data. Note that previous pupillometry studies using luminance flicker have tended to fit sine-waves to the data, rather than using Fourier analysis (e.g. Spitschan et al., 2014). The Fourier approach is more robust to noise at other frequencies and has been used in some previous studies (see Barrionuevo et al., 2014; Barrionuevo and Cao, 2016). Additionally, it makes the pupillometry analysis consistent with standard practice in steady-state EEG analysis (e.g. Figueira et al., 2022).

@@

We fitted a computational model to the data from Experiments 1 \& 3 using a hierarchical Bayesian approach. The model behaviour is displayed in Figure \ref{fig:modelfigure}e-h, with empirical data replotted in Figure \ref{fig:modelfigure}a-d for comparison. In general, the model captures the key characteristics of the empirical data. Note that there are some minor discrepancies, which are a consequence of the hierarchical nature of the modelling. In brief, the model is fitted to the amplitudes for each participant, and group-level parameter estimates are derived based on these fits (see Table \ref{tab:paramtable}). This procedure discards the phase information, whereas the empirical averages are coherently averaged across participants (retaining phase information). This explains the amplitude differences between model and data, particularly at low target contrast levels, but are of no consequence for the pattern of relative responses across conditions.

We were particularly interested in comparing the weight of interocular suppression across data sets. We therefore plot the posterior distributions for this parameter for all four data sets (see Figure \ref{fig:modelfigure}i). The key finding is that the pupillometry results (green distribution) display a much greater weight of interocular suppression compared with the other data sets (grey, purple and yellow distributions). There is no overlap between the pupillometry distribution and any of the other three. All four distributions are also meaningfully below a weight of 1 -- the value that previous work using grating stimuli would predict (Baker and Wade, 2017; Meese et al., 2006). These results offer an explanation of the empirical data: the strong interocular suppression for the pupillometry data is consistent with the weak binocular facilitation, and measurable dichoptic masking observed using that method. The weaker suppression for the other experiments is consistent with the near-linear binocular facilitation effects, and absence of dichoptic masking.

@@

Using a novel paradigm that combines EEG and pupillometry, we found surprising results for the binocular combination of flickering light. In the visual cortex, the processing of temporal luminance modulations seems to happen approximately linearly, with no evidence of interocular suppression. In the subcortical pathway, signal combination is more non-linear, with evidence of interocular suppression. This pattern of results was also supported by computational modelling, which showed a much greater suppressive weight for the pupillometry data compared to the EEG data. Additionally, we found that perception of flickering light is consistent with a near-linear summation process, consistent with the cortical (EEG) responses.

The results of our main experiment were unexpected for both the pupillometry and the EEG measures. Previous studies investigating binocular combination of spatial patterns (i.e.~sine wave grating stimuli) are all consistent with strong interocular suppression and weak binocular facilitation (Baker and Wade, 2017; Meese et al., 2006; Moradi and Heeger, 2009). Our second experiment ruled out the possibility that these differences were due to the lower temporal frequency (2Hz) used here. However, there is evidence of more extensive binocular facilitation for a range of other stimuli. For example, Quaia et al. (2018) observed a strong binocular facilitation (or `supersummation') in the reflexive eye movement response to rapidly moving stimuli, and Spitschan and Cajochen (2019) report a similar result in archival data on melatonin suppression due to light exposure. In the auditory system, interaural suppression of amplitude modulation also appears to be weak when measured using a similar steady-state paradigm (Baker et al., 2020). Finally, psychophysical matching experiments using static stimuli also show near-linear behaviour for luminance increments (Anstis and Ho, 1998; Baker et al., 2012; Levelt, 1965), though not for luminance decrements (Anstis and Ho, 1998). Overall, this suggests that strong interocular normalization may be specific to spatial pattern vision, and not a general feature of sensory signal combination.

@@

@@

We have demonstrated that binocular combination of flickering light differs between cortical and subcortical pathways. Flicker was also associated with substantially weaker interocular suppression, and stronger binocular facilitation, compared to combination of spatial luminance modulation in visual cortex. Our computational framework for understanding signal combination permits direct comparisons between disparate experimental paradigms and data types. We anticipate that this will help elucidate the constraints the brain faces when combining different types of signals to govern perception, action and biological function.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Barrionuevo2016}{}}%
Barrionuevo PA, Cao D. 2016. Luminance and chromatic signals interact differently with melanopsin activation to control the pupil light response. \emph{J Vis} \textbf{16}:29. doi:\href{https://doi.org/10.1167/16.11.29}{10.1167/16.11.29}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Carpenter2017}{}}%
Carpenter B, Gelman A, Hoffman MD, Lee D, Goodrich B, Betancourt M, Brubaker M, Guo J, Li P, Riddell A. 2017. Stan: A probabilistic programming language. \emph{Journal of Statistical Software, Articles} \textbf{76}:1--32. doi:\href{https://doi.org/10.18637/jss.v076.i01}{10.18637/jss.v076.i01}

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Kassner2014}{}}%
Kassner M, Patera W, Bulling A. 2014. Pupil: An open source platform for pervasive eye tracking and mobile gaze-based interactionProceedings of the 2014 {ACM} International Joint Conference on Pervasive and Ubiquitous Computing: Adjunct Publication. UbiComp '14 Adjunct; {ACM}. pp. 1151--1160. doi:\href{https://doi.org/10.1145/2638728.2641695}{10.1145/2638728.2641695}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-McDougal2008}{}}%
McDougal DH, Gamlin PD. 2008. Pupillary control pathwaysThe Senses: A Comprehensive Reference. Elsevier. pp. 521--536.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Moradi2009}{}}%
Moradi F, Heeger DJ. 2009. Inter-ocular contrast normalization in human visual cortex. \emph{J Vis} \textbf{9}:13.1--22. doi:\href{https://doi.org/10.1167/9.3.13}{10.1167/9.3.13}

\leavevmode\vadjust pre{\hypertarget{ref-Norcia2015}{}}%
Norcia AM, Appelbaum LG, Ales JM, Cottereau BR, Rossion B. 2015. The steady-state visual evoked potential in vision research: A review. \emph{J Vis} \textbf{15}:4. doi:\href{https://doi.org/10.1167/15.6.4}{10.1167/15.6.4}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Quaia2018}{}}%
Quaia C, Optican LM, Cumming BG. 2018. Binocular summation for reflexive eye movements. \emph{J Vis} \textbf{18}:7. doi:\href{https://doi.org/10.1167/18.4.7}{10.1167/18.4.7}

@@

>>>>

\title{Different rules for binocular combination of luminance flicker in cortical and subcortical pathways}
\author{Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade \& Daniel H. Baker}
\date{2023-01-07}

@@

How does the brain combine information across the eyes? In primary visual cortex, spatial patterns are subject to `ocularity invariance', where monocular and binocular responses are rendered equal by mutual suppression between the eyes. Here we asked whether this invariance holds for pure changes in luminance. We collected steady-state EEG and pupillometry data simultaneously, and find strong deviations from ocularity invariance both in the cortex and also in the subcortical pathway that determines pupil diameter. In cortex, we find strong binocular facilitation, and negligible interocular suppression from cortical sources, whereas measurements of pupil diameter showed weaker facilitation and stronger suppression. This was not purely a consequence of the temporal stimulus parameters; pronounced binocular facilitation was also observed at faster flicker rates. Near-linear binocular combination was also found for the same stimuli using a perceptual matching task. A hierarchical Bayesian implementation of a standard binocular combination model confirms that interocular suppression was substantially weaker for our EEG and matching data compared with the pupillometry results. These findings illustrate that ocularity invariance is not a ubiquitous feature of visual processing, and that the brain can repurpose a generic algorithm for different functions by adjusting parameters such as the weight of suppression between pathways.

@@

The brain must combine information across multiple sensory inputs to derive a coherent percept of the external world. This involves a process of signal combination both within (Baker and Wade, 2017) and between (Ernst and Banks, 2002) the senses. Binocular vision is a useful test case for signal combination, as the inputs to the two eyes overlap substantially (in species with forward-facing eyes), and the neural locus is well-established (Hubel and Wiesel, 1962). Much of our knowledge about binocular combination derives from studies on the contrast response of the `canonical' visual pathway, in which signals pass from the eyes to primary visual cortex (V1), via the lateral geniculate nucleus (LGN) (Purves et al., 2008). However, signals are also combined across the eyes in the network of subcortical nuclei that govern pupil diameter in response to absolute light levels (McDougal and Gamlin, 2008) and much less is known about the computations that operate in these subcortical pathways. In addition, our primary purpose here is to investigate the computations governing signal combination in these two anatomically distinct pathways in response to luminance changes.

For pattern vision, binocular presentation confers greater sensitivity to low contrast targets than monocular presentation. This is known as binocular summation, with summation ratios (the relative improvement under binocular presentation) at detection threshold lying between \(\sqrt{2}\) and 2 (Baker et al., 2018; Campbell and Green, 1965). This advantage is lost at high stimulus intensities, where both psychophysical performance (contrast discrimination thresholds) (Legge, 1984; Meese et al., 2006) and neural activity (Baker and Wade, 2017; Moradi and Heeger, 2009) are approximately equal for monocular and binocular presentation. Contemporary models of binocular vision (Ding and Sperling, 2006; Meese et al., 2006) advocate a process of interocular suppression that normalizes the two eyes inputs at high contrasts and negates the binocular advantage. This is consistent with our everyday experience of `ocularity invariance' (Baker et al., 2007): perceived contrast does not change when one eye is opened and closed.

The pupillary light reflex is an automatic constriction of the iris sphincter muscles in response to increases in light levels, which causes the pupil to shrink (McDougal and Gamlin, 2008). There is a clear binocular component to this reflex, as stimulation of one eye still causes constriction of the other eye's pupil (termed the consensual response; Wyatt and Musselman (1981)). Importantly, the neuroanatomical pathway involved completely bypasses the canonical cortical pathway (retina to V1), instead involving a network of subcortical nuclei, including the Pretectal Olivary nucleus, Superior Cervical ganglion, and Edinger-Westphal nucleus (Angée et al., 2021; Mathôt, 2018; McDougal and Gamlin, 2008; Wang and Munoz, 2015). To account for the consensual response, these brain regions must combine information from the left and right eyes (Doesschate and Alpern, 1967), yet the computation that achieves this is achieved is unclear.

To address this shortcoming, we designed an experiment that allowed us to simultaneously record electrophysiological and pupillometric responses to monocular and binocular stimuli. This novel paradigm allowed us to probe both cortical (using EEG) and subcortical (using a binocular eyetracker) pathways simultaneously in response to flickering light, and make quantitative comparisons between them. Periodic flicker entrains both neural (Norcia et al., 2015) and pupil (Spitschan et al., 2014) responses at the flicker frequency, enabling precise estimation of response amplitudes in the Fourier domain. We followed up our main experiment with additional exploration of the effect of stimulus frequency, and a psychophysical matching experiment measuring perceived flicker intensity (i.e.~temporal contrast). The results are interpreted using a hierarchical Bayesian computational model of binocular vision, and reveal that subcortical pathways implement stronger interocular suppression than the canonical cortical pathway.

@@

The stimuli were two discs of achromatic flickering light with a diameter of 3.74 degrees, presented on a black background. The same stimuli were used for all three experiments. Four dark red lines were added around both discs to help with their fusion into one binocular disc (see insert in Figure \ref{fig:pupildata}b for an example of the fused stimulus). The discs were viewed through a four-mirror stereoscope, which used front silvered mirrors to avoid internal reflections, and meant that participants saw a single fused disc. The use of a stereoscope allowed us to modulate the stimuli in three different ocular configurations: monocular, binocular, and dichoptic. Note that during monocular presentation of flicker, the unstimulated eye still saw the static (non-flickering) disc of mean luminance.

@@

The experiment was conducted in a windowless room, in which the only light source was the monitor. The participants sat at 99 cm from the monitor and the total optical viewing distance (through the stereoscope) was 107 cm. The experiment was carried out in a single session lasting 45 minutes in total, divided in three blocks of 15 minutes each. In each block, there were 60 trials lasting 15 seconds each (12s of stimulus presentation, with an interstimulus interval of 3s). The participants were given no task other than to look at the fixation cross in the middle of the disc while trying to minimise their blinking during the presentation period.

We included six distinct ocular conditions, each at five temporal contrast levels (combined factorially) relative to the mean luminance: 6, 12, 24, 48 and 96\%. Contrast was defined as temporal Michelson contrast; the difference between maximum and minimum luminances, scaled by the mean and expressed as a percentage. In the first three conditions, the discs flickered at 2 Hz, in either a monocular, binocular, or dichoptic arrangement. In the dichoptic condition the non-target eye saw a fixed contrast of 48\%. In the remaining three conditions (termed the cross-frequency conditions) one eye's disc flickered at 1.6Hz, and the other eye's disc flickered at 2Hz. We also tested monocular (one eye sees 1.6Hz flicker, the other sees mean luminance), binocular (one eye sees each frequency at the target contrast) and dichoptic (target stimulus flickering at 2Hz, mask contrast of 48\% at 1.6Hz in the other eye) arrangements. The rationale for flickering both eyes at 2Hz is that we can then measure summation behaviour between the eyes in the pupil and EEG response at 2Hz. The rationale for flickering the eyes at different frequencies is that this permits measurement of suppression between the eyes (i.e.~the reduction in the 2Hz response when a 1.6Hz mask component is added to the other eye). We counterbalanced presentation of the target stimulus across the left and right eyes.

@@

EEG data were converted from the ANT-EEProbe format to a compressed csv text file using a custom Matlab script and components of the EEGlab toolbox (Delorme and Makeig, 2004). The data for each participant were then loaded into R for analysis, where a ten-second waveform for each trial at each electrode was extracted (omitting the first two seconds). The Fourier transform of each waveform was calculated, and the complex spectrum stored in a matrix. All repetitions of each condition were then averaged for each electrode. They were then averaged across four occipital electrodes (\emph{POz}, \emph{Oz}, \emph{O1}, \emph{O2}), to obtain individual results. Finally, these were averaged across participants to obtain the group results. All averaging was performed in the complex domain and therefore retained the phase information (i.e.~coherent averaging), and at each stage we excluded data points with a Mahalanobis distance exceeding \(D\) = 3 from the complex-valued mean (see Baker, 2021). For statistical comparisons of complex-valued data, we use the \(ANOVA^2_{circ}\) statistic described by Baker (2021). This is a multivariate extension of ANOVA that assumes equal variance of the real and imaginary Fourier components, or equivalently, an extension of the \(T^2_{circ}\) statistic of Victor and Mast (1991) that can compare more than two conditions.

A similar analysis pipeline was adopted for the pupillometry data. The data were converted from mp4 videos to a csv text file using the Pupil Player software (Kassner et al., 2014), which estimated pupil diameter for each eye on each frame using a 3D model of the eyeball. The individual data were then loaded into R for analysis, where again a ten-second waveform for each trial in each eye was extracted (excluding the first two seconds after stimulus onset). We interpolated across any dropped or missing frames to ensure regular and continuous sampling over time. The Fourier transform was calculated for each waveform, and all repetitions of each condition were pooled across eye and then averaged. Finally, data were averaged across all participants to obtain the group results. Again, we used coherent averaging, and excluded outlying data points in the same way as for the EEG data. Note that previous pupillometry studies using luminance flicker have tended to fit a single sine-wave at the fundamental frequency, rather than using Fourier analysis (e.g. Spitschan et al., 2014). The Fourier approach is more robust to noise at other frequencies and has been used in some previous studies (see Barrionuevo et al., 2014; Barrionuevo and Cao, 2016). Additionally, it makes the pupillometry analysis consistent with standard practice in steady-state EEG analysis (e.g. Figueira et al., 2022).

@@

We fitted a computational model to the data from Experiments 1 \& 3 using a hierarchical Bayesian approach. The model behaviour is displayed in Figure \ref{fig:modelfigure}e-h, with empirical data replotted in Figure \ref{fig:modelfigure}a-d for comparison. In general, the model captures the key characteristics of the empirical data. Note that there are some minor discrepancies, which are a consequence of the hierarchical nature of the modelling. In brief, the model is fitted to the amplitudes for each participant, and group-level parameter estimates are derived based on these fits (see Table \ref{tab:paramtable}). This procedure discards the phase information, whereas the empirical averages are coherently averaged across participants (retaining phase information). This explains the amplitude differences between model and data, particularly at low target contrast levels, but is of little consequence for the pattern of relative responses across conditions, which is our main focus here.

We were particularly interested in comparing the weight of interocular suppression across data sets. We therefore plot the posterior distributions for this parameter for all four data sets (see Figure \ref{fig:modelfigure}i). The key finding is that the pupillometry results (green distribution) display a much greater weight of interocular suppression compared with the other data sets (grey, purple and yellow distributions). There is no overlap between the pupillometry distribution and any of the other three. All four distributions are also meaningfully below a weight of 1 -- the value that previous work using grating stimuli would predict (Baker and Wade, 2017; Meese et al., 2006), and the peak location of our prior distribution (black curve). These results offer an explanation of the empirical data: the strong interocular suppression for the pupillometry data is consistent with the weak binocular facilitation, and measurable dichoptic masking observed using that method. The weaker suppression for the other experiments is consistent with the near-linear binocular facilitation effects, and absent dichoptic masking.

@@

Using a novel paradigm that combines EEG and pupillometry, we found surprising results for the binocular integration of flickering light. In the visual cortex, the processing of spatially-uniform temporal luminance modulations seems to happen approximately linearly, with no evidence of interocular suppression. In the subcortical pathway, signal combination is more non-linear, with evidence of interocular suppression. This pattern of results was confirmed by computational modelling, which showed a much greater suppressive weight for the pupillometry data compared to the EEG data. Additionally, we found that perception of flickering light is consistent with a near-linear summation process, consistent with the cortical (EEG) responses.

The results of our main experiment were unexpected for both the pupillometry and the EEG measures. Previous studies investigating binocular combination of spatial patterns (i.e.~sine wave grating stimuli) are all consistent with strong interocular suppression and weak binocular facilitation at high contrasts (Baker and Wade, 2017; Meese et al., 2006; Moradi and Heeger, 2009). Our second experiment ruled out the possibility that these differences were due to the lower temporal frequency (2Hz) used here. However, there is evidence of more extensive binocular facilitation for a range of other stimuli. For example, Quaia et al. (2018) observed a strong binocular facilitation (or `supersummation') in the reflexive eye movement response to rapidly moving stimuli, and Spitschan and Cajochen (2019) report a similar result in archival data on melatonin suppression due to light exposure. In the auditory system, interaural suppression of amplitude modulation also appears to be weak when measured using a similar steady-state paradigm (Baker et al., 2020). Finally, psychophysical matching experiments using static stimuli also show near-linear behaviour for luminance increments (Anstis and Ho, 1998; Baker et al., 2012; Levelt, 1965), though not for luminance decrements (Anstis and Ho, 1998). Overall, this suggests that strong interocular normalization may be specific to spatial pattern vision, and not a general feature of sensory signal combination.

@@

Classic studies investigating the neurophysiological architecture of V1 reported that cells in cytochrome-oxidase `blobs' (Horton and Hubel, 1981; Livingstone and Hubel, 1984) are biased towards low spatial frequencies (Edwards et al., 1995; Tootell et al., 1988), and relatively insensitive to stimulus orientation (Horton and Hubel, 1981; Livingstone and Hubel, 1984; though see Economides et al., 2011). As the blob regions are embedded within ocular dominance columns (Horton and Hubel, 1981), they are also largely monocular (Livingstone and Hubel, 1984; Tychsen et al., 2004). More recent work has reported psychophysical evidence for unoriented chromatic (Gheiratmand et al., 2013) and achromatic (Meese and Baker, 2011) mechanisms, that also appear to be monocular. Our use of luminance flicker might preferentially stimulate these mechanisms, perhaps explaining why our EEG data show little evidence of binocular interactions. Indeed, our EEG results could potentially be explained by a model involving entirely non-interacting monocular channels, with the binocular facilitation effects we find (e.g.~Figures \ref{fig:BSratios} \& \ref{fig:TFdata}) owing to additivity of the electrophysiological response across independent monocular cells, rather than requiring binocular neurons. However our matching data (Figure \ref{fig:matchingdata}), as well as everyday perceptual experience, indicates that luminance signals must be combined physiologically at some stage, and that this process still involves weak or absent interocular suppression.

@@

We have demonstrated that binocular combination of flickering light differs between cortical and subcortical pathways. Flicker was also associated with substantially weaker interocular suppression, and stronger binocular facilitation, compared to combination of spatial luminance modulations in visual cortex. Our computational framework for understanding signal combination permits direct comparisons between disparate experimental paradigms and data types. We anticipate that this will help elucidate the constraints the brain faces when combining different types of signals to govern perception, action and biological function.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

Supported by BBSRC grant BB/V007580/1 awarded to DHB and ARW, and Wellcome Trust grant 213616/Z/18/Z to AB.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Barrionuevo2016}{}}%
Barrionuevo PA, Cao D. 2016. Luminance and chromatic signals interact differently with melanopsin activation to control the pupil light response. \emph{J Vis} \textbf{16(11)}:29. doi:\href{https://doi.org/10.1167/16.11.29}{10.1167/16.11.29}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Carpenter2017}{}}%
Carpenter B, Gelman A, Hoffman MD, Lee D, Goodrich B, Betancourt M, Brubaker M, Guo J, Li P, Riddell A. 2017. Stan: A probabilistic programming language. \emph{Journal of Statistical Software} \textbf{76}:1--32. doi:\href{https://doi.org/10.18637/jss.v076.i01}{10.18637/jss.v076.i01}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Economides2011}{}}%
Economides JR, Sincich LC, Adams DL, Horton JC. 2011. Orientation tuning of cytochrome oxidase patches in macaque primary visual cortex. \emph{Nat Neurosci} \textbf{14}:1574--80. doi:\href{https://doi.org/10.1038/nn.2958}{10.1038/nn.2958}

\leavevmode\vadjust pre{\hypertarget{ref-Edwards1995}{}}%
Edwards DP, Purpura KP, Kaplan E. 1995. Contrast sensitivity and spatial frequency response of primate cortical neurons in and around the cytochrome oxidase blobs. \emph{Vision Res} \textbf{35}:1501--23. doi:\href{https://doi.org/10.1016/0042-6989(94)00253-i}{10.1016/0042-6989(94)00253-i}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Gheiratmand2013}{}}%
Gheiratmand M, Meese TS, Mullen KT. 2013. Blobs versus bars: Psychophysical evidence supports two types of orientation response in human color vision. \emph{Journal of Vision} \textbf{13(1)}:2. doi:\href{https://doi.org/10.1167/13.1.2}{10.1167/13.1.2}

\leavevmode\vadjust pre{\hypertarget{ref-Horton1981}{}}%
Horton JC, Hubel DH. 1981. Regular patchy distribution of cytochrome oxidase staining in primary visual cortex of macaque monkey. \emph{Nature} \textbf{292}:762--4. doi:\href{https://doi.org/10.1038/292762a0}{10.1038/292762a0}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Kassner2014}{}}%
Kassner M, Patera W, Bulling A. 2014. Pupil: An open source platform for pervasive eye tracking and mobile gaze-based interaction.Proceedings of the 2014 {ACM} International Joint Conference on Pervasive and Ubiquitous Computing: Adjunct Publication. UbiComp '14 Adjunct; {ACM}. pp. 1151--1160. doi:\href{https://doi.org/10.1145/2638728.2641695}{10.1145/2638728.2641695}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Livingstone1984}{}}%
Livingstone MS, Hubel DH. 1984. Anatomy and physiology of a color system in the primate visual cortex. \emph{J Neurosci} \textbf{4}:309--56. doi:\href{https://doi.org/10.1523/JNEUROSCI.04-01-00309.1984}{10.1523/JNEUROSCI.04-01-00309.1984}

@@

\leavevmode\vadjust pre{\hypertarget{ref-McDougal2008}{}}%
McDougal DH, Gamlin PD. 2008. Pupillary control pathways.The Senses: A Comprehensive Reference. Elsevier. pp. 521--536.

\leavevmode\vadjust pre{\hypertarget{ref-Meese2011}{}}%
Meese TS, Baker DH. 2011. A reevaluation of achromatic spatio-temporal vision: Nonoriented filters are monocular, they adapt, and can be used for decision making at high flicker speeds. \emph{{iPerception}} \textbf{2}:159--82. doi:\href{https://doi.org/10.1068/i0416}{10.1068/i0416}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Moradi2009}{}}%
Moradi F, Heeger DJ. 2009. Inter-ocular contrast normalization in human visual cortex. \emph{J Vis} \textbf{9(3)}:13. doi:\href{https://doi.org/10.1167/9.3.13}{10.1167/9.3.13}

\leavevmode\vadjust pre{\hypertarget{ref-Norcia2015}{}}%
Norcia AM, Appelbaum LG, Ales JM, Cottereau BR, Rossion B. 2015. The steady-state visual evoked potential in vision research: A review. \emph{J Vis} \textbf{15(6)}:4. doi:\href{https://doi.org/10.1167/15.6.4}{10.1167/15.6.4}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Quaia2018}{}}%
Quaia C, Optican LM, Cumming BG. 2018. Binocular summation for reflexive eye movements. \emph{J Vis} \textbf{18(4)}:7. doi:\href{https://doi.org/10.1167/18.4.7}{10.1167/18.4.7}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Tootell1988}{}}%
Tootell RB, Silverman MS, Hamilton SL, Switkes E, De Valois RL. 1988. Functional anatomy of macaque striate cortex. V. Spatial frequency. \emph{J Neurosci} \textbf{8}:1610--24. doi:\href{https://doi.org/10.1523/JNEUROSCI.08-05-01610.1988}{10.1523/JNEUROSCI.08-05-01610.1988}

\leavevmode\vadjust pre{\hypertarget{ref-Tychsen2004}{}}%
Tychsen L, Wong AM-F, Burkhalter A. 2004. Paucity of horizontal connections for binocular vision in V1 of naturally strabismic macaques: Cytochrome oxidase compartment specificity. \emph{The Journal of Comparative Neurology} \textbf{474}:261--275. doi:\href{https://doi.org/10.1002/cne.20113}{10.1002/cne.20113}
